Measuring Directionality in Double Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors

### 2013-07-22 Direction_v6.tex
Version 6 of draft. Comments included from emails on July 22. --Christoph

\section{Liquid Scintillator Detectors}
Liquid scintillators are `cocktails' of aromatic hydrocarbons. When
charged particles move through a scintillator, the molecules are
excited, predominantly via the non-localized electrons in the
$\pi$-bonds of the phenyl groups\cite{birks_book}. Vibrational
and rotational modes of the molecules are turned into heat within
picoseconds through collisions with other molecules.  Within $\sim$10
picoseconds, the $\pi$-electrons de-excite to the first excited state
from higher levels through radiation-less transitions. The first
excited state can de-excite through photon emission. There are two
characteristic times for this de-excitation, depending if the singlet
state or the triplet state was excited.  The singlet state will
de-excite within nanoseconds while the triplet state de-excites on the
order of 10's or 100's of nanoseconds. These two processes are
fluorescence and phosphorescence respectively. The exact time
constants for these processes are determined by the composition of the
scintillator.

The absorption and emission spectra overlap at some level
in all molecules. Consequently, if there is only one type of molecule in
the scintillator cocktail the light output is reduced due to
inefficiencies in the energy transfer through multiple absorption and
re-emission processes. Aromatic solutes or fluorophores are added to the
primary solvent to shift the wavelengths of the photons to higher values 
where the scintillator is more transparent. This
wavelength-shifting is also used to match the quantum efficiency as a
function of wavelength for the photodetectors being used. One typical
scintillator mixture uses pseudocumene as the solvent with 1-5~g/L of
PPO as the fluorophore. This mixture has a peak emission at about 400~nm
where bialkali photomultiplier tubes (PMTs) are most sensitive and the
pseudocumene is relatively transparent.

@@

In $0\nu\beta\beta$, the electrons emerge with a combined energy equal
to the Q-value of the particular isotope. The individual
electrons follow distributions of energies and angular correlations, 
a probable case being equal division of energy between back-to-back
electrons\cite{phasespace}. This case is shown in FIG.~\ref{detector_view}
for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
half-life is inversely proportional to the phase-space factor, isotopes with
higher Q-values are preferred and due to backgrounds from the daughters
of the $^{238}$U and $^{232}$Th decay chains those with Q-values at or
above 2.6~MeV are most often considered for $0\nu\beta\beta$ searches.
There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb},
but only a handful with large Q-values.  Most of the high Q-value candidates
have been considered as a dopant for a liquid scintillator:
$^{150}$Nd(Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr(Q=3.350~MeV)\cite{zr1},
$^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=3.350~MeV)\cite{qdot},
$^{116}$Cd(Q=2.802~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller},
$^{136}$Xe(Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn(Q=2.29~MeV)\cite{sn1}.
Xenon gas readily dissolves into liquid scintillator. For the other isotopes,
a suitable organometallic compound needs to be found that produces a stable
scintillator with a long attenuation length in the wavelength region of
interest. As an alternative to single-atom-doping, recently nanocrystals 
formed by candidate isotopes have been explored as a dopant\cite{qdot,qdot2}.

\section{Geant4 Simulation}
In order to study the effects relevant to directional reconstruction
in liquid scintillators, a
\GEANT\cite{geant4one,geant4two} simulation has been constructed. The
simulation uses \GEANT~Version 4.9.6 with the default liquid scintillator
optical model, in which optical photons are
assigned the group velocity in the wavelength region of normal
dispersion.

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
        \caption[]{Two back-to-back electrons with 1.41~MeV each
        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay) at
        the center of the sphere with initial directions along the x
        and -x-axis. Only Cherenkov photons (green lines) are drawn to illustrate the
        directionality of the event. \label{detector_view}}
        \end{center}
\end{figure}

The detector geometry is a sphere of 6.5~m radius filled with
scintillator. FIG.~\ref{detector_view} shows the geometry and the
Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
default scintillator properties have been chosen to match the KamLAND
scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene
(1,2,4-trimethylbenzene) and 1.52~g/l PPO (2,5-diphenyloxazole). The
scintillator properties implemented in the simulation include the
atomic composition and density ($\rho$ = 0.78~g/ml), the
wavelength-dependent attenuation length\cite{tajimaMaster} and
refractive index\cite{OlegThesis}, the scintillation emission
spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns)
and emission decay time constants ($\tau_{d1}$ = 6.9~ns and
$\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 
0.13)\cite{tajimaThesis}, scintillator light yield (9030 photons/MeV)
and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}. 
Variations from the baseline KamLAND case are
discussed below. Re-emission of absorbed photons in the scintillator
bulk volume and scattering have not yet been included, but are not
expected to change the conclusions here.

@@

\begin{figure*}[tbh]
\begin{center}
        \subfigure[ ~Default
        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
        \subfigure[ ~Increased TTS
        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
        \subfigure[ ~Red-sensitive
        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
        \caption[]{Photo-electron (PE) arrival times after application
        of the transit-time spread (TTS) for the simulation
        of 1000 electrons (5~MeV) with different values of the TTS and
        wavelength response. PEs from
        Cherenkov light (black, solid line) and scintillation light
        (red, dotted line) are
        compared. The dash-dotted vertical line illustrates a time cut at
        34~ns. (a) Default simulation: bialkali photocathode and TTS =
        0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
        from scintillation and 108~PEs from Cherenkov light. (b)
        Default simulation settings except for TTS = 1.28~ns (KamLAND
        17 in. PMTs). After the 34.0~ns time cut we get 349~PEs from
        scintillation and 88~PEs from Cherenkov light. (c) Default
        simulation settings except for a GaAsP photocathode. After the
        34.0~ns time cut we get 226~PEs from scintillation and 229~PEs
        from Cherenkov light. \label{time_plots_comparison}}
        \end{center}
\end{figure*}

@@

We first discuss results for the default simulation settings described
in the previous section. FIG.~\ref{time_plots_comparison} (a) shows
the TTS-smeared photoelectron (PE) detection times for 1000 simulated
electrons with 5~MeV energy in the center of the detector with initial
momentum directions coinciding with the x-axis. The photoelectrons
induced by Cherenkov light arrive earlier, as expected due to the
instantaneous emission and the higher average photon speed compared to
scintillation light. There is however significant overlap of the two
arrival time distributions.

@@

FIG.~\ref{Cherenkov_cone} displays the angular distribution of PE
hits after the time cut. Although this time cut is a simplification of actual time
reconstruction effects, we can use it to indicate the spatial
distribution of hits in the early time window. The Cherenkov ring structure
can be clearly seen in the peak near 46\textdegree, demonstrating
that the directional signal conveyed by the Cherenkov photons is not
erased by scattering of the initial 5~MeV electrons.

When the 17-inch KamLAND PMTs\cite{tajimaMaster,kume_1983} (TTS =
1.28~ns) are used in the simulation, the broadening of the time
distributions leads to a strongly decreased ratio of Cherenkov over
scintillation light ($R_{C/S}=0.25$) for $t<34$~ns (see
FIG.~\ref{time_plots_comparison} (b)). This shows that a low
photodetector TTS is critical for directionality reconstruction and
motivates the use of novel detector types.

\section{Detector Wavelength Response}
\label{detector_wavelength_response_sec} 
In addition to decreasing the photodetector TTS to enhance $R_{C/S}$,
it is possible to optimize the wavelength-dependence of the
photocathode. Since Cherenkov photons which passed through meters of
scintillator have on average longer wavelengths than scintillation
photons, a photodetector which is more sensitive at long wavelengths
increases not only the absolute number of PEs but also the ratio
between Cherenkov- and scintillation-induced PEs.

We have run the simulation with the QE of an extended red-sensitive
GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}.
FIG.~\ref{time_plots_comparison}(c) shows the results for the
modified simulation with high QE in the red spectral region. The
higher absolute number of photoelectrons coming from Cherenkov light
(factor of $\approx$ 2) and the increased Cherenkov/scintillation
ratio ($R_{C/S}=1.6$) in the early time window would significantly improve the
directionality reconstruction.

@@

The emission spectrum of commercial alloyed core/shell
CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured in
Ref.\cite{qdot2}. This spectrum shows a symmetric peak centered
around 461~nm with FWHM = 29~nm.  In order to isolate the effect of
the different emission spectrum, the other simulation settings,
including the KamLAND absorption spectrum, were kept unchanged; we
find $R_{C/S}=0.17$ for the default 34~ns timing cut.  Compared to
the default case shown in FIG.~\ref{time_plots_comparison}(a) the
separation is worse (as expected) because the scintillation light
wavelengths are higher than in the KamLAND emission spectrum.

@@

\begin{figure*}[tbh]
        \begin{center}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
        \subfigure{\includegraphics[scale=0.4]{graphs/hDir_34ns.pdf}}
        \subfigure{\includegraphics[scale=0.4]{graphs/hVtx_34ns.pdf}}
        \caption[]{\label{fig:reco} (Left) The reconstructed direction,
        $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|},
        \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons
        (5~MeV). In the simulation the electrons are produced along the
        x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
        from the center of the 6.5m-radius detector, $\vec{r}$ =
        (0,0,0). Only photons with arrival time of $t<$ 34~ns are used
        in the reconstruction. The quantum efficiency of the bialkali
        photocathode is taken into account. (Right) The reconstructed
        vertex position, $(x,y,z)$, for the same simulation.}

@@

The timing studies show that in the early time window, $t\leq$
34.0~ns, the ratio $R_{C/S}$ is high. In this paper, we apply
reconstruction tools for a water Cherenkov detector, WCSimAnalysis,
to the problem of reconstructing the position and direction of 5~MeV
electrons from this early light.  WCSimAnalysis is a water Cherenkov
reconstruction package developed for the Long Baseline Neutrino
Experiment (LBNE collaboration)\cite{Blake}. It provides a framework
for generic event cleaning, track reconstruction, and particle
identification, and comes equipped with variety of pre-built
algorithms. It is continuing to be expanded using new track-fitting
techniques for water Cherenkov detectors\cite{Sanchez2012525} based on
advanced photosensors with sub-cm imaging capabilities and timing
resolutions below 100 picoseconds\cite{LAPPDSum,LAPPDTDR}.

@@

For the purpose of testing the reconstruction algorithm we use 1000
simulated electrons with an energy of 5~MeV. The electrons are
simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). FIG.~\ref{fig:reco}
shows the vertex reconstruction. The vertex is reasonably well
reconstructed around the center of the detector, $\vec{r}$ = (0,0,0),
except along the x-axis. The RMS values of the distributions for all three
reconstructed coordinates are smaller than 3.5~cm. The shift along
the x-axis is due to two effects for which the reconstruction has to
use average values rather than the unknown true value for each
event: the wavelength and hence the speed of the light in the medium,
and the point of emission for each of the photons, reconstructed
as coming from a common point. The reconstruction of the direction
also is shown in Fig.~\ref{fig:reco}. It shows that for the majority
of the events the initial electron direction is reconstructed well.
This is a promising result given the simplicity of the algorithms
used.

\section{Conclusions}
The ability to reconstruct direction in kiloton-scale scintillation
detectors would be a major technological advance for neutrino
experiments, especially those also
searching for neutrino-less double-beta decay. More generally, this technique could
be applied wherever scintillation-based detectors are used. A \GEANT~simulation of a simple spherical detector corresponding to a kiloton
of scintillator shows that timing on the order of 0.1~ns is required
to separate the directional Cherenkov light from the more abundant
scintillation light. This separation can be improved using
photodetectors with more red sensitivity and liquid scintillators with
a more narrow emission spectrum shifted to shorter
wavelengths. Furthermore, simple reconstruction algorithms adapted
from those for water Cherenkov detectors are able to reliably
reconstruct the position and direction of 5~MeV electrons. More
detailed simulation and advanced reconstruction algorithms will need
to be developed to move to lower energies and more complicated event
topography, such as those in neutrino-less double-beta decay, but the technique already
appears promising.

>>>>

\section{Liquid Scintillator Detectors}
Liquid scintillators are `cocktails' of aromatic hydrocarbons. When
charged particles move through a scintillator, the molecules are
excited, predominantly via the non-localized electrons in the
$\pi$-bonds of the phenyl groups\cite{birks_book}. Vibrational
and rotational modes of the molecules are turned into heat within
picoseconds through collisions with other molecules.  Within $\sim$10
picoseconds, the $\pi$-electrons de-excite to the first excited state
from higher levels through radiationless transitions. The first
excited state can de-excite through photon emission. There are two
characteristic times for this de-excitation, depending if the singlet
state or the triplet state was excited.  The singlet state will
de-excite within nanoseconds while the triplet state de-excites on the
order of 10's or 100's of nanoseconds. These two processes are
fluorescence and phosphorescence respectively. The exact time
constants for these processes are determined by the composition of the
scintillator.

The absorption and emission spectra overlap at some level
in all molecules. Consequently, if there is only one type of molecule in
the scintillator cocktail the light output is reduced due to
inefficiencies in the energy transfer through multiple absorption and
re-emission processes. Aromatic solutes or fluorophores are added to the
primary solvent to shift the wavelengths of the photons to higher values 
where the scintillator is more transparent. This
wavelength-shifting is also used to match the quantum efficiency as a
function of wavelength for the photodetectors being used. One typical
scintillator mixture uses pseudocumene as the solvent with 1-5~g/l of
PPO as the fluorophore. This mixture has a peak emission at about 400~nm
where bialkali photomultiplier tubes (PMTs) are most sensitive and the
pseudocumene is relatively transparent.

@@

In $0\nu\beta\beta$, the electrons emerge with a combined energy equal
to the Q-value of the particular isotope. The individual
electrons follow distributions of energies and angular correlations, 
a probable case being equal division of energy between back-to-back
electrons\cite{phasespace}. This case is shown in Fig.~\ref{detector_view}
for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
half-life is inversely proportional to the phase-space factor, isotopes with
higher Q-values are preferred and due to backgrounds from the daughters
of the $^{238}$U and $^{232}$Th decay chains those with Q-values at or
above 2.6~MeV are most often considered for $0\nu\beta\beta$ searches.
There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb},
but only a handful with large Q-values.

Most of the high Q-value candidates
have been considered as a dopant for a liquid scintillator:
$^{150}$Nd(Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr(Q=3.350~MeV)\cite{zr1},
$^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=3.350~MeV)\cite{qdot},
$^{116}$Cd(Q=2.802~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller},
$^{136}$Xe(Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn(Q=2.29~MeV)\cite{sn1}.
Xenon gas readily dissolves into liquid scintillator. For the other isotopes,
a suitable organometallic compound needs to be found that produces a stable
scintillator with a long attenuation length in the wavelength region of
interest. As an alternative to single-atom-doping, recently nanocrystals 
formed by candidate isotopes have been explored as a dopant\cite{qdot,qdot2}.

\section{Geant4 Simulation}
In order to study the effects relevant to directional reconstruction
in liquid scintillators, a
\GEANT\cite{geant4one,geant4two} simulation has been constructed. The
simulation uses \GEANT~version 4.9.6 with the default liquid scintillator
optical model, in which optical photons are
assigned the group velocity in the wavelength region of normal
dispersion.

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
        \caption[]{The detector geometry and coordinate system.
        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.41~MeV each
        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay). The electrons originate at
        the center of the sphere with initial directions along the x
        and -x-axis. Only Cherenkov photons are drawn to illustrate the
        directionality of the event. \label{detector_view}}
        \end{center}
\end{figure}

The detector geometry is a sphere of 6.5~m radius filled with
scintillator. Figure~\ref{detector_view} shows the geometry and the
Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
default scintillator properties have been chosen to match the KamLAND
scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene
(1,2,4-trimethylbenzene) and 1.52~g/l PPO (2,5-diphenyloxazole). The
scintillator properties implemented in the simulation include the
atomic composition and density ($\rho$ = 0.78~g/ml), the
wavelength-dependent attenuation length\cite{tajimaMaster} and
refractive index\cite{OlegThesis}, the scintillation emission
spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns)
and emission decay time constants ($\tau_{d1}$ = 6.9~ns and
$\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 
0.13)\cite{tajimaThesis}, scintillator light yield (9030 photons/MeV)
and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}. 
Variations from the baseline KamLAND case are
discussed below. Re-emission of absorbed photons in the scintillator
bulk volume and scattering have not yet been included, but are not
expected to change the conclusions here.

@@

\begin{figure*}[tbh]
\begin{center}
        \subfigure[ ~Default
        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
        \subfigure[ ~Increased TTS
        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
        \subfigure[ ~Red-sensitive
        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
        \caption[]{Photo-electron (PE) arrival times after application
        of the transit-time spread (TTS) for the simulation
        of 1000 electrons (5~MeV) with different values of the TTS and
        wavelength response. PEs from
        Cherenkov light (black, solid line) and scintillation light
        (red, dotted line) are
        compared. The dash-dotted vertical line illustrates a time cut at
        34.0~ns. (a) Default simulation: bialkali photocathode and TTS =
        0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
        from scintillation and 108~PEs from Cherenkov light. (b)
        Default simulation settings except for TTS = 1.28~ns (KamLAND
        17 in. PMTs). After the 34.0~ns time cut we get 349~PEs from
        scintillation and 88~PEs from Cherenkov light. (c) Default
        simulation settings except for a GaAsP photocathode. After the
        34.0~ns time cut we get 226~PEs from scintillation and 229~PEs
        from Cherenkov light. \label{time_plots_comparison}}
        \end{center}
\end{figure*}

@@

We first discuss results for the default simulation settings described
in the previous section. Figure~\ref{time_plots_comparison} (a) shows
the TTS-smeared photoelectron (PE) detection times for 1000 simulated
electrons with 5~MeV energy in the center of the detector, with initial
momentum directions coinciding with the x-axis. The photoelectrons
induced by Cherenkov light arrive earlier, as expected due to the
instantaneous emission and the higher average photon speed compared to
scintillation light. There is, however, significant overlap of the two
arrival time distributions.

@@

Figure~\ref{Cherenkov_cone} displays the angular distribution of PE
hits after the time cut. Although this time cut is a simplification of actual time
reconstruction effects, we can use it to indicate the spatial
distribution of hits in the early time window. The Cherenkov ring structure
can be clearly seen in the peak near 46\textdegree, demonstrating
that the directional signal conveyed by the Cherenkov photons is not
erased by scattering of the initial 5~MeV electrons.

When the 17-inch KamLAND PMTs\cite{tajimaMaster,kume_1983} (TTS =
1.28~ns) are used in the simulation, the broadening of the time
distributions leads to a strongly decreased ratio of Cherenkov over
scintillation light ($R_{C/S}=0.25$) for $t<34.0$~ns (see
Fig.~\ref{time_plots_comparison} (b)). This shows that a low
photodetector TTS is critical for directionality reconstruction and
motivates the use of novel photodetector types.

\section{Detector Wavelength Response}
\label{detector_wavelength_response_sec} 
In addition to decreasing the photodetector TTS to enhance $R_{C/S}$,
it is possible to optimize the wavelength-dependence of the
photocathode. Since Cherenkov photons which pass through meters of
scintillator have on average longer wavelengths than scintillation
photons, a photodetector which is more sensitive at long wavelengths
increases not only the absolute number of PEs but also the ratio
between Cherenkov- and scintillation-induced PEs.

We have run the simulation with the QE of an extended red-sensitive
GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}.
Figure~\ref{time_plots_comparison}(c) shows the results for the
modified simulation with high QE in the red spectral region. The
higher absolute number of photoelectrons coming from Cherenkov light
(factor of $\approx$ 2) and the increased Cherenkov/scintillation
ratio ($R_{C/S}=1.6$) in the early time window would significantly improve the
directionality reconstruction.

@@

The emission spectrum of commercial alloyed core/shell
CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured in
Ref.\cite{qdot2}. This spectrum shows a symmetric peak centered
around 461~nm with FWHM = 29~nm.  In order to isolate the effect of
the different emission spectrum, the other simulation settings,
including the KamLAND absorption spectrum, were kept unchanged; we
find $R_{C/S}=0.17$ for the default 34.0~ns timing cut.  Compared to
the default case shown in Fig.~\ref{time_plots_comparison}(a) the
separation is worse (as expected) because the scintillation light
wavelengths are longer than in the KamLAND emission spectrum.

@@

\begin{figure*}[tbh]
        \begin{center}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
        \subfigure{\includegraphics[scale=0.4]{graphs/hDir_34ns.pdf}}
        \subfigure{\includegraphics[scale=0.4]{graphs/hVtx_34ns.pdf}}
        \caption[]{\label{fig:reco} (Left) The reconstructed direction,
        $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|},
        \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons
        (5~MeV). In the simulation the electrons are produced along the
        x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
        from the center of the 6.5m-radius detector, $\vec{r}$ =
        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
        in the reconstruction. The quantum efficiency of the bialkali
        photocathode is taken into account. (Right) The reconstructed
        vertex position, $(x,y,z)$, for the same simulation.}

@@

The timing studies show that in the early time window, $t\leq$
34.0~ns, the ratio $R_{C/S}$ is high, improving the event selection. In this section, we apply
reconstruction tools for a water Cherenkov detector, WCSimAnalysis,
to the problem of reconstructing the position and direction of 5~MeV
electrons from this early light.  WCSimAnalysis is a water Cherenkov
reconstruction package developed for the Long Baseline Neutrino
Experiment (LBNE collaboration)\cite{Blake}. It provides a framework
for generic event cleaning, track reconstruction, and particle
identification, and comes equipped with variety of pre-built
algorithms. It is continuing to be expanded using new track-fitting
techniques for water Cherenkov detectors\cite{Sanchez2012525} based on
advanced photosensors with sub-cm imaging capabilities and timing
resolutions below 100 picoseconds\cite{LAPPDSum,LAPPDTDR}.

@@

For the purpose of testing the reconstruction algorithm we use 1000
simulated electrons with an energy of 5~MeV. The electrons are
simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). Figure~\ref{fig:reco}
shows the vertex reconstruction. The vertex is reasonably well
reconstructed around the center of the detector, $\vec{r}$ = (0,0,0),
except along the x-axis. The RMS values of the distributions for all three
reconstructed coordinates are smaller than 3.5~cm. The shift along
the x-axis is due to two effects for which the reconstruction has to
use average values rather than the unknown true value for each
event: the wavelength and hence the speed of the light in the medium,
and the point of emission for each of the photons, reconstructed
as coming from a common point. The reconstruction of the direction
also is shown in Fig.~\ref{fig:reco}. It shows that for the majority
of the events the initial electron direction is reconstructed well.
This is a promising result given the simplicity of the algorithms
used.

\section{Conclusions}
The ability to reconstruct direction in kiloton-scale scintillation
detectors would be a major technological advance for neutrino
experiments, especially those also
searching for neutrino-less double-beta decay. More generally, this technique could
be applied wherever scintillation-based detectors are used. A \GEANT~simulation of a simple spherical detector corresponding to a kiloton
of scintillator shows that timing on the order of 0.1~ns is required
to separate the directional Cherenkov light from the more abundant
scintillation light. The separation can be improved using
photodetectors with more red sensitivity and liquid scintillators with
a more narrow emission spectrum shifted to shorter
wavelengths. Furthermore, simple reconstruction algorithms adapted
from those for water Cherenkov detectors are able to reliably
reconstruct the position and direction of 5~MeV electrons. More
detailed simulation and advanced reconstruction algorithms will need
to be developed to move to lower energies and more complicated event
topography, such as those in neutrino-less double-beta decay, but the technique already
appears promising.
